Market-based Mechanisms for Resource Scheduling

### 2013-12-18 paper/paper.tex
Typo fix
Typo fixes

Large-scale computing clusters offer the potential for greater efficiency through
pooling of computational resources, but must contend with the problem of
allocating these resources to users with competitive and potentially
incompatible demands. Such conflicts are typically arbitrated by some \emph{fairness}
criterions, such as dominant resource fairness (DRF) \cite{drf}, which benefit from their
simplicity, efficiency, and non-manipulability. Unfortunately, when applied in
realistic settings --- when dealing with inhomogeneous resources, time-varying
demands or deadlines, requirements for data locality, etc.---typical implementations
of these systems cease to be either
efficient or non-manipulable.

@@

We stress the distinction between this use of price signals and the more typical
motivation. We are most interested in price signals as an \emph{abstraction} which
programs can use to express their constraints, rather than as a mechanism to
extract revenue for the cluster operator or to arbitrate between competing
external interests. To consider a simple case, suppose there are two programs
that are contending for a shared resource $R$. Each could either make use of $R$ or
some alternative resource $R'$ (for example, a different machine in the cluster at
a different time). Our proposal is to have each program articulate the value of
resource $R$ (compared to the benchmark of not receiving resource $R$, and then
contending for resource $R'$ at a later date). The primary alternatives that have
been considered are:
\vspace{-4mm}
\begin{enumerate}
\itemsep0em
  \item Ignore the relative value of resource $R$ and $R'$, and make allocation
    decisions entirely on the basis of fairness considerations (for example by
    ensuring that each framework has an equal number of every resource). This is
    the approach taken by Mesos, although it in fact achieves better behavior
    precisely by \emph{violating} the strategy-proofness of its underlying allocation
    policy, and providing incentives to frameworks to strategically decline a
    resource in order to receive a better allocation in the future.
  \item Understand the entire optimization problem faced by each problem, and
    make explicit judgments of the relative value of $R$ and $R'$ within the
    scheduler. This violates the spirit of Mesos, and in general appears to be
    difficult given the complexity and scale of the cluster scheduling problem.
\end{enumerate}

@@

There have bene many ad hoc proposals to address some of these difficulties.
For example, Jockey is a system which accomodates deadlines in a fair-share
scheduler by increasing the allotted shares for users which are appraoching
a deadline.
However, these techniques tend to violate the theoretical guarantees
of the underlying systems more starkly, and to require extensive tuning
in order to achieve good performance.
We seek a more general framework for addressing these difficulties.

@@

A naive representation of this function would use an exponential amount of
space. While we can consider several methods greatly varying in
complexity for encoding this function more compactly, we use the following
two-level representation:
\begin{itemize}
  \item A list of \emph{tasks}, with the assumption that a framework's
    willingness to pay for any set of tasks is simply the total
    willingness to pay for tasks in that set.
  \item For each task a list of possible resources which can be used to perform
    that task, together with a willingness to pay for that set of resources.
\end{itemize}
While this obviously cannot represent all possible $\mathcal{B}$, it still
allows enough flexibility to cover all of the applications we allude to
in the paper---to handle dealdines, inhomogeneous resources, fragmentation,
tasks of differing priorities, and so on.

\subsection{Ascending-bid clock auction}
Once the master has collected a set of bids, it uses an ascending-bid clock
auction to determine which of the bids to fulfill, and how much it should charge
for each of the tasks which win the auction. The auction consists of a number of
rounds which continue until no changes occur in the final round. To avoid the
need for communication between the master and the frameworks after each round,
we use a \emph{proxy auction} where each framework makes a single bid for all
rounds and agents running with the master represent each of them within the
master.

@@

Formally, each framework calculates the current net present value
of their accumulated wealth together with their current obligations
and stream of income, and then bids a price proportional
to this total. This guarantees that the frameworks will
maintain roughly equal budgets (and hence use roughly equal shares
of the cluster), since any framework which became richer
than the others would begin winning auctions.
\vspace{-5mm}
\paragraph{Adjusting bids} To modify the 
default policy to accomodate inhomogeneous resources,
times, or tasks, we can simply modify framework's
bids to adjust for the desirability of the resource being
purchased or the importance of the purpose they are being used for.

For example, if a task will take less time when run on a machine
which already has relevant data loaded into memory,
a framework can increase its willingness to pay for resources on that machine
by an amount depending on the anticipated speedup.
%XXX mention why we don't care about any properties other than those that we mention
\vspace{-5mm}
\paragraph{Defensive bidding} One negative consequence of using a closed economy,
in which the frameworks bid using currency which is entirely internal to the cluster,
%XXX mention this distinction elsewhere, and the fact that we can do either
is the possibility of \emph{hoarding}, in which one framework
declines to use any resources during a certain period so that it can amass
a much larger budget than any other resources.
It can then spend this budget at a future point in order to monopolize the cluster,
which may have undesirable conseuqences for frameworks that need uninterrupted service.

@@

By the definition of our auction, the price $p_B$ exceeds user A's willingness to pay for resources $R_B$ 
(or else user A's proxy would have continued bidding and the auction would not have concluded), 
while the price $p_A$ is less than user A's willingness to pay for resources $R_A$ 
(or else user A's proxy would have stopped bidding earlier). 
In particular, user A prefers their package to user B's package. 
Since at the start of the round user A preferred their allocation at their price to user B's allocation at user B's price, 
this property also holds at the end of the round. 
In this step we rely on our assumption that user A's valuations for their own holdings do not change after the auction is announced 
(otherwise it may be possible for user A to become envious in between rounds as more information is revealed, rather than during a round of the auction). 
We note in passing that similar assumptions are required to prove that DRF is envy-free in the dynamic setting,
even in a simple rather than realistic enviornment.
\qed
\end{pproof}
%XXX insert discussion of strategy-proofness of DRF, which mentions (1) the fact that you need to ignore complexities of items / trust users to incorporate them into their bids, (2) the fact that you need to ignore the consequences of your behavior for other participants
\vspace{-8mm}
\begin{theorem}[Local strategy-proofness] 
Assuming that users are indifferent to changes in each others' budgets,
and that users' willingness to pay 
incorporates accurate expectations of effects on future allocations, 
the optimal behavior in our auction is to honestly declare willingness to pay.
\end{theorem}
\vspace{-7mm}
\begin{pproof}
We give the standard argument for the dominant strategy truthfulness (strategy-proofness) of Vickrey auctions. 
Indeed, suppose on the contrary 
that $B_1$ and $B_2$ are two different bids for user A; that $R_1$ and $R_2$ are the resources received after making either of these bids and $p_1$, $p_2$ are the associated prices; that $B_1$ is an honest bid; and that user A prefers $(R_2, p_2)$ to $(R_1, p_1)$.

@@

\subsection{Grid computing}
Job schedulers like TORQUE \cite{torque}, Condor \cite{condor}, and LSF
\cite{lsf} have long been used for solving the problem of resource allocation
within a high-performance-computing (HPC) or grid-computing environment.
For the most part, they do not use a two-level scheduling model like we do but
instead operate as a queue for predefined jobs which are run as resources become
available. Machines and job requirements also tend to show more homogeniety
within such environments (for example, the ratio between CPU, memory, and disk),
so sthe amount of optimization possible with pricing is lower.

@@

\subsection{Resource allocation}
Dominant resource fairness~\cite{drf} is an allocation policy whereby the user with the smallest dominant share is next allocated resources. The user's dominant share in a multi-resource environment is the largest share the user possesses of any of the cluster's various resources.

>>>>

Large-scale computing clusters offer the potential for greater efficiency through
pooling of computational resources, but must contend with the problem of
allocating these resources to users with competitive and potentially
incompatible demands. Such conflicts are typically arbitrated by some \emph{fairness}
criterion, such as dominant resource fairness (DRF) \cite{drf}, which benefit from their
simplicity, efficiency, and non-manipulability. Unfortunately, when applied in
realistic settings --- when dealing with inhomogeneous resources, time-varying
demands or deadlines, requirements for data locality, etc.---typical implementations
of these systems cease to be either
efficient or non-manipulable.

@@

We stress the distinction between this use of price signals and the more typical
motivation. We are most interested in price signals as an \emph{abstraction} which
programs can use to express their constraints, rather than as a mechanism to
extract revenue for the cluster operator or to arbitrate between competing
external interests. To consider a simple case, suppose there are two programs
that are contending for a shared resource $R$. Each could either make use of $R$ or
some alternative resource $R'$ (for example, a different machine in the cluster at
a different time). Our proposal is to have each program articulate the value of
resource $R$ (compared to the benchmark of not receiving resource $R$, and then
contending for resource $R'$ at a later date). The primary alternatives that have
been considered are:
\vspace{-4mm}
\begin{enumerate}
\itemsep0em
  \item Ignore the relative value of resource $R$ and $R'$, and make allocation
    decisions entirely on the basis of fairness considerations (for example by
    ensuring that each framework has an equal number of every resource). This is
    the approach taken by Mesos, although it in fact achieves better behavior
    precisely by \emph{violating} the strategy-proofness of its underlying allocation
    policy, and providing incentives to frameworks to strategically decline a
    resource in order to receive a better allocation in the future.
  \item Understand the entire optimization problem faced by each problem, and
    make explicit judgements of the relative value of $R$ and $R'$ within the
    scheduler. This violates the spirit of Mesos, and in general appears to be
    difficult given the complexity and scale of the cluster scheduling problem.
\end{enumerate}

@@

There have been many ad hoc proposals to address some of these difficulties.
For example, Jockey is a system which accommodates deadlines in a fair-share
scheduler by increasing the allotted shares for users which are approaching
a deadline.
However, these techniques tend to violate the theoretical guarantees
of the underlying systems more starkly, and to require extensive tuning
in order to achieve good performance.
We seek a more general framework for addressing these difficulties.

@@

A naive representation of this function would use an exponential amount of
space. While we can consider several methods greatly varying in
complexity for encoding this function more compactly, we use the following
two-level representation:
\begin{itemize}
  \item A list of \emph{tasks}, with the assumption that a framework's
    willingness to pay for any set of tasks is simply the total
    willingness to pay for tasks in that set.
  \item For each task a list of possible resources which can be used to perform
    that task, together with a willingness to pay for that set of resources.
\end{itemize}
While this obviously cannot represent all possible $\mathcal{B}$, it still
allows enough flexibility to cover all of the applications we allude to
in the paper---to handle deadlines, inhomogeneous resources, fragmentation,
tasks of differing priorities, and so on.

\subsection{Ascending-bid clock auction}
Once the master has collected a set of bids, it uses an ascending-bid clock
auction to determine which of the bids to fulfil, and how much it should charge
for each of the tasks which win the auction. The auction consists of a number of
rounds which continue until no changes occur in the final round. To avoid the
need for communication between the master and the frameworks after each round,
we use a \emph{proxy auction} where each framework makes a single bid for all
rounds and agents running with the master represent each of them within the
master.

@@

Formally, each framework calculates the current net present value
of their accumulated wealth together with their current obligations
and stream of income, and then bids a price proportional
to this total. This guarantees that the frameworks will
maintain roughly equal budgets (and hence use roughly equal shares
of the cluster), since any framework which became richer
than the others would begin winning auctions.
\vspace{-5mm}
\paragraph{Adjusting bids} To modify the 
default policy to accommodate inhomogeneous resources,
times, or tasks, we can simply modify framework's
bids to adjust for the desirability of the resource being
purchased or the importance of the purpose they are being used for.

For example, if a task will take less time when run on a machine
which already has relevant data loaded into memory,
a framework can increase its willingness to pay for resources on that machine
by an amount depending on the anticipated speedup.
%XXX mention why we don't care about any properties other than those that we mention
\vspace{-5mm}
\paragraph{Defensive bidding} One negative consequence of using a closed economy,
in which the frameworks bid using currency which is entirely internal to the cluster,
%XXX mention this distinction elsewhere, and the fact that we can do either
is the possibility of \emph{hoarding}, in which one framework
declines to use any resources during a certain period so that it can amass
a much larger budget than any other resources.
It can then spend this budget at a future point in order to monopolize the cluster,
which may have undesirable consequences for frameworks that need uninterrupted service.

@@

By the definition of our auction, the price $p_B$ exceeds user A's willingness to pay for resources $R_B$ 
(or else user A's proxy would have continued bidding and the auction would not have concluded), 
while the price $p_A$ is less than user A's willingness to pay for resources $R_A$ 
(or else user A's proxy would have stopped bidding earlier). 
In particular, user A prefers their package to user B's package. 
Since at the start of the round user A preferred their allocation at their price to user B's allocation at user B's price, 
this property also holds at the end of the round. 
In this step we rely on our assumption that user A's valuations for their own holdings do not change after the auction is announced 
(otherwise it may be possible for user A to become envious in between rounds as more information is revealed, rather than during a round of the auction). 
We note in passing that similar assumptions are required to prove that DRF is envy-free in the dynamic setting,
even in a simple rather than realistic environment.
\qed
\end{pproof}
%XXX insert discussion of strategy-proofness of DRF, which mentions (1) the fact that you need to ignore complexities of items / trust users to incorporate them into their bids, (2) the fact that you need to ignore the consequences of your behavior for other participants
\vspace{-8mm}
\begin{theorem}[Local strategy-proofness] 
Assuming that users are indifferent to changes in each others' budgets,
and that users' willingness to pay 
incorporates accurate expectations of effects on future allocations, 
the optimal behavior in our auction is to honestly declare willingness to pay.
\end{theorem}
\vspace{-7mm}
\begin{pproof}
We give the standard argument for the dominant strategy truthfulness (strategy-proofness) of Vickrey auctions. 
Indeed, suppose on the contrary 
that $B_1$ and $B_2$ are two different bids for user A; that $R_1$ and $R_2$ are the resources received after making either of these bids and $p_1$, $p_2$ are the associated prices; that $B_1$ is an honest bid; and that user A prefers $(R_2, p_2)$ to $(R_1, p_1)$.

@@

\subsection{Grid computing}
Job schedulers like TORQUE \cite{torque} or Condor \cite{condor} have long been used for solving the problem of resource allocation
within a high-performance-computing (HPC) or grid-computing environment.
For the most part, they do not use a two-level scheduling model like we do but
instead operate as a queue for predefined jobs which are run as resources become
available. Machines and job requirements also tend to show more homogeneity
within such environments (for example, the ratio between CPU, memory, and disk),
so the amount of optimization possible with pricing is lower.

@@

\subsection{Resource allocation}
Dominant resource fairness~\cite{drf} is an allocation policy whereby the user
with the smallest dominant share is next allocated resources. The user's
dominant share in a multi-resource environment is the largest share the user
possesses of any of the cluster's various resources. The Mesos cluster manager
relies on the DRF policy to decide which framework to next offer resources.
Section \ref{sec:theory} contains a more theoretical discussion of DRF and 
its benefits compared to a market-based mechanism.
